--- 
title: "A Minimal Book Example"
author: "Yihui Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
# bibliography: [book.bib, packages.bib]
biblio-style: apalike2
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Introduction {-}

Should I be using linear regression, logistics regression, SVM, neural networks or one of at least a dozen other statistical/machine learning algorithms when trying to model my data?  This is the fundamental question which has driven the authorship of this book.  The problem is that the answer to that question for the most part is always the same, it depends.  The answer depends not only on the data itself, but also why you are modeling in the first place, i.e., what question are you trying to answer with your model, as well as a whole host of other factors.  The purpose of this book is to help you identify the important factors for choosing a model,  given those factors which model(s) you should choose and why, and then finally how to implement that model in R or python.

# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

For now, you have to install the development versions of **bookdown** from Github:

```{r eval=FALSE}
devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need to install XeLaTeX.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


