---
title: "R Bootcamp"
author: "Michoel Snow"
date: "July 2, 2017"
fontsize: 30pt
output:
  beamer_presentation:
#    keep_tex: true
#    toc: true
    slide_level: 3
    includes:
      in_header: beamerstyle.txt
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# The Basics
## Getting R up and Running
### Installing R
If you don't have them already installed you need to download and install R and R studio  

- [Download R](https://cran.r-project.org/)  
- [R Studio](https://www.rstudio.com/products/rstudio/download/)

### Introduction

- R is an open source programming environment
- It is most commonly used for statistical computing and graphics
- The basic functionality of R can be extended with 'packages'

<!-- ### Sheena Easton and Game Theory -->

<!-- Sheena Easton describes the following scenario for her baby: -->

<!-- 1. Takes the morning train -->
<!-- 2. Works from nine 'til five -->
<!-- 3. Takes another train home again -->
<!-- 4. Finds Sheena Easton waiting for him -->


<!-- - Akin to Holmes-Moriarty game (see: von Neumann and Morgenstern) -->


<!-- ## Never Gonna Give You Up -->

<!-- ### Rick Astley's Re-election Platform -->

<!-- Rick Astley's campaign promises: -->

<!-- - Never gonna give you up -->
<!-- - Never gonna let you down -->
<!-- - Never gonna run around and desert you -->
<!-- - Never gonna make you cry -->
<!-- - Never gonna say goodbye -->
<!-- - Never gonna tell a lie and hurt you. -->

<!-- ## R Markdown -->

<!-- This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. -->

<!-- When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. -->

<!-- ### Slide with Bullets -->

<!-- $X^2$ -->

<!-- - Bullet 1 -->
<!-- - Bullet 2 -->
<!-- - Bullet 3 -->

### Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

### Slide with Plot

```{r pressure}
plot(pressure)
```

